---
title: "Example install packages"
author: "Anja Bråthen Kristoffersen"
date: "14 7 2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

If you want to get access to packages that haven't been released on CRAN, you have to edit your .Rprofile. To do this run 
 
```{r}
usethis::edit_r_profile()
```


To use packgaes from FHI write:

options(repos=structure(c(
  FHI="https://folkehelseinstituttet.github.io/drat/",
  CRAN="https://cran.rstudio.com"
)))

in the .Rprofile file.

NB make sure that " is written correctly. Might change when you copy.

Then write install.packages("LIBRARY NAME"), see an example with som packages you will need to use below:

```{r}


list.of.packages <- c("org", "fs", "data.table", "ggplot2", "gridExtra", "grid", "flowCore", "flowSOM", "betareg", "ComplexHeatmap", "Rmarkdown")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) {
  install.packages(new.packages)
}


```


