---
title: "Clean up gating"
author: "Anja Bråthen Kristoffersen"
date: "16 2 2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document describes the different lines of the file "example clean up gating alle filer.R". To perform your own gating save the file "example clean up gating alle filer.R" from github folkehelseinstituttet/cytof (finn sti) give it a new name and change the necessary paths and values throughout the document to use it on your own data. 


Define a variable data_path that contains the path of the folder where all your uncleaned files are saved. I prefer using the function path from the library fs. When using a function from a given library the way to get the function is by library_name::function_name so in this case fs::path 


```{r}
data_path <- fs::path("C:", "CyToF data", "immun_aga", "panel1 RAA")
```

which gives the path

```{r}
data_path
```

Now define the paths to the folder where you want save 1) the new files that are clean up gated and 2) figures of the gating performance. Do also define the path where the files script used are saved.
```{r}
outDataPath <- fs::path("C:", "CyToF data", "immun_aga", "panel 1 cleaned data")
out_result <- fs::path("C:", "CyToF data", "immun_aga", "figures panel 1 cleaning")
scriptPath <- fs::path("C:", "CyToF data", "fra github")
```

I have already made many functions that can be used for reading, transforming, gating and plotting in R. This functions can be found on github/SJEKK ADRESSE. Save the files gating_functions.R, ploting_functions, read_data_function.R and transformation_functions.R into your script folder and source them into R so that you might use them later


```{r}
source(fs::path(scriptPath, "gating_functions.R"))
source(fs::path(scriptPath, "ploting_functions.R"))
source(fs::path(scriptPath, "read_data_functions.R"))
source(fs::path(scriptPath, "transformation_functions.R"))
```

file.info() and list.files() are functions in R that can be  used to find the name of all files in one folder 

```{r}
fcs_files <- fs::path(data_path, rownames(file.info(list.files(data_path))))
```

to see which files you want to read into R:

```{r}
fcs_files
```

basename is a function that give you the name of the file without the path. By using the function grepl() we choose only the files that contains .fcs. 

setwd() is a function who set the working directory. Here we want the working directory to be where the uncleand cytof files are. dirname() gives the directroy/folder of one file, and fcs_files[1] gives the first file in the vector fcs_files.

```{r}
files_to_open <- basename(fcs_files)
files_to_open <- files_to_open[grepl(".fcs", files_to_open)]
setwd(dirname(fcs_files[1]))
```

The variable file_names are made shorter (without ".fcs") by using the function gsub, the function gsub first takes the pattern that you want to substitute and then the pattern you want to replace it with in the vector which is the last variable in the function call. This is done to have shorter unique names on each dataset/file for later use.

```{r}
file_names <- gsub(".fcs", "", files_to_open)
# Read the files into a flowset
```

the number of .fcs files in data_path are found by the function length()

```{r}
n_files <- length(file_names)
```

A vector filenumber containing 1, 2, 3, .... to n_files are made for later use

```{r}
filenumbers <- 1:n_files
filenumbers
```

Two empty matrixes to take care of how many percentage of each file that are lost by each gating and in total of all previos gating are defined as percent_lost_each_gating and percent_lost_from_full_dataset, respectivly. These data.frames are given the column name of the gating and the rownames of the files that were gated.

```{r}
percent_lost_each_gating <- as.data.frame(matrix(NA, ncol = 9, nrow = n_files))
percent_lost_from_full_dataset <-   as.data.frame(matrix(NA, ncol = 9, nrow = n_files))

colnames(percent_lost_each_gating) <- c("Ce140Di", "Residual", "Center", "Offset", "Width",
                                        "Event_length", "Pt194Di", "Ir191Di", "Ir193Di")

colnames(percent_lost_from_full_dataset) <- c("Ce140Di", "Residual", "Center", "Offset", "Width",
                                              "Event_length", "Pt194Di", "Ir191Di", "Ir193Di")

rownames(percent_lost_each_gating) <- file_names
rownames(percent_lost_from_full_dataset) <- file_names

```



This is a function that has to be made to the dataset that are used. It tells how to plot the files you use. Here I have 6 files and I want to plot them with 3 rows and 2 columns. I want to save the file in the variable filename, and plot_list is a list that I will produce later that contain a plot per file. To write comments into an R code you use # in the beginning of the line. I have here written a verson of this function that plots 66 plots in 6 rows and 11 columns.

```{r}
plotSignal <- function(plot_list){
  g <- gridExtra::grid.arrange(plot_list[[1]], plot_list[[2]], plot_list[[3]], plot_list[[4]], plot_list[[5]],
                          plot_list[[6]], ncol = 3, nrow = 2)
  return(g)
}

# plotTiff <- function(plot_list, filnavn){
#   tiff(fs::path(out_result, filnavn), height = 1000, width = 1800)
#   gridExtra::grid.arrange(plot_list[[1]], plot_list[[2]], plot_list[[3]], plot_list[[4]], plot_list[[5]],
#                           plot_list[[6]],  plot_list[[7]], plot_list[[8]], plot_list[[9]], plot_list[[10]],
#                           plot_list[[11]], plot_list[[12]], plot_list[[13]], plot_list[[14]], plot_list[[15]],
#                           plot_list[[16]],  plot_list[[17]], plot_list[[18]], plot_list[[19]], plot_list[[20]],
#                           plot_list[[21]], plot_list[[22]], plot_list[[23]], plot_list[[24]], plot_list[[25]],
#                           plot_list[[26]],  plot_list[[27]], plot_list[[28]], plot_list[[29]], plot_list[[30]],
#                           plot_list[[31]], plot_list[[32]], plot_list[[33]], plot_list[[34]], plot_list[[35]],
#                           plot_list[[36]],  plot_list[[37]], plot_list[[38]], plot_list[[39]], plot_list[[40]],
#                           plot_list[[41]], plot_list[[42]], plot_list[[43]], plot_list[[44]], plot_list[[45]],
#                           plot_list[[46]],  plot_list[[47]], plot_list[[48]], plot_list[[49]], plot_list[[50]],
#                           plot_list[[51]], plot_list[[52]], plot_list[[53]], plot_list[[54]], plot_list[[55]],
#                           plot_list[[56]],  plot_list[[57]], plot_list[[58]], plot_list[[59]], plot_list[[60]],
#                           plot_list[[61]], plot_list[[62]], plot_list[[63]], plot_list[[64]], plot_list[[65]],
#                           plot_list[[66]],  ncol = 11, nrow = 6)
#    dev.off()
# }


```



```{r}
# If because of memory problems the data cleaning has to be done in groups  this is an exampel of how to replace i below. (if using for-loop it can not be devided into script parts as below but rather on big unit as in example clean up gating alle filer.R)
#for(ii in 1:floor(n_files/6)){
#  
#  i <- filenumber[((ii-1) * 6 + (1:6))]
#  i <- i[!is.na(i)]
```

You could either look at all files in the same time or devide it into blocks of files. Here is an example with all files. If memory problems it is possible to divide the files to clean into groups using a loop. 

i is here the files to  use
```{r}
i <- 1:length(file_names)

ii <- 1
```

Read all files in data_path into a list object fcs_data

```{r}
fcs_data_with_info <- read_some_data_from_folder(data_path, file_number = i)
  fcs_data <- fcs_data_with_info$fcs_data
  file_names <- factor(fcs_data_with_info$file_names, levels = fcs_data_with_info$file_names)
  rm(fcs_data_with_info)
```

By using the function get_params_fcs_data() the parameters of the first element of the list fcs_data are obtained.

```{r}
params <- get_params_fcs_data(fcs_data[[1]])
params
```

## Beads gating

beads channels can be "Ce140Di", "Eu151Di", "Eu153Di", "Ho165Di" or "Lu175Di" or maybe something else in your dataset. Here we use grepl to get the correct name used in params. the string "140|151|153|165|175" tells grepl to look for 140, 151, 153, 165 or 175 in the column name of the data.frame params. R is case sensitive so it is important to write the name correctly.

```{r}
#***************************************************
  #gating on Beads ----
  #***************************************************
  
  
  bead_channels <- as.character(params$name[grepl("140|151|153|165|175", params$name)])
bead_channels
```
here we choose to do the beads gating on Ce140Di, first making the list beads_data where I have arc_sinh transformed the signal data

```{r}
x <- "Ce140Di"
beads_data <-  arc_sinh_transform_selected_channels(fcs_data = fcs_data, channels = bead_channels)
```

by using the function number_of_events_raw_data() we get a vector with number of events in the dataset

```{r}
number_of_events_raw_data <-  number_of_events(data = fcs_data, file_names = file_names)
number_of_events_raw_data
```

since it takes time and memory to plot some many events we here choose to randomly plot 10000 events per file. This events are found by the function random_events() which gives a list of as many elements that you have files and that contains a vector with 10000 numbers representing the events to plot. 

```{r}
random_events_for_plotting <- random_events(number_of_events_raw_data, n = 10000)
```

When performing beads gating the perpose is to exclude events with high signal. 

Here the function  find_gate_perc_height_upper_noise() takes the dataset that you want to make the gates based on, here beads_data in column channel x, and a value upper_perc_height, here 0.001, the proportion height of the max of the density curve. The max of the density curve is here shown with the red line. while 0.001 times this height is shown with the green line. The gate should be placed where the green horizontal line cross the density curve. Here shown by the green vertical line. The function find_gate_perc_height_upper_noise() take a list of files and return a vector in the same length as the list, i.e. one gate per file. 

```{r, echo  = F, width = 12, height =5}
#example only in word document to show how the function work.
signal <- beads_data[[1]][, x]
dens <- density(signal)
posible <- dens$y > max(dens$y) * 0.001
xposible <- which(posible[2:length(posible)] - posible[1:(length(posible)-1)] == -1)[1]
value <- dens$x[xposible]
y_max <- max(dens$y)
x_max <- dens$x[which(dens$y == y_max)]


plot(dens, lwd =3)
abline(h = y_max, col = "red", lwd = 2)
abline(h = dens$y[xposible], col = "green")
lines(c(value, value), c(0,1), col ="green", lwd = 2)


##NB forandr navn på funksjon til prop ikke perc


```






```{r}
#find upper_gate for noise gating based on Ce140Di
  upper_gate_Ce140Di <- find_gate_perc_height_upper_noise(data = beads_data, channel = x, upper_perc_height = 0.001)
```

make a time signal plot where the signal is plotted against time and the gate is shown. 

```{r}
time_signal_plots <- time_signal_plot(data = beads_data, random_events = random_events_for_plotting, 
                                        channel = x,  plot_title = file_names, upper_gate = upper_gate_Ce140Di)
  
time_signal_plots[1] # to see first plot
```

All the plots are plotted together in a tiff file. If to many files in one plot the plotTiff function could be changed.

```{r}
plotSignal(plot_list = time_signal_plots)
```

The gate could also be plotted in a density plot made by the function density_plot()

```{r}
density_plots <- density_plot(data = beads_data, channel = x, plot_title = file_names, upper_gate = upper_gate_Ce140Di)
  #density_plots # to see the plots
  
  
  # tiff(fs::path(outFigPath, paste0("fig1_bead_gating", ii, ".tiff")))
   density_plots
  # dev.off()
```


Before continuing with the script it should be decided if the gate is ok  or something should be adjusted. 
If ok, proced using the gate and adjust the dataset by excluding the events with signal above upper_gate_Ce140Di.

* The function events_to_keep_after_gating() find which events to keep per dataset. 
* The function update_data_based_on_events_to_keep() update the fcs_data
* The number of events per file after gating are found by number_of_events()
* The percentages lost by this gating and for full dataset is calculated and saved in the matrices percent_lost_from_full_dataset and percent_lost_from_full_dataset
* A new vector of random_events_for_plotting is made by the function random_events()

```{r}
events_to_keep_after_gating <- events_to_keep(data = beads_data, channel = "Ce140Di",  upper_gate = upper_gate_Ce140Di)
  
  #overwrite the raw  dataset (will take lot of space if we make one new each time, and do not need it)
  fcs_data <- update_data_based_on_events_to_keep(data = fcs_data, kept_events = events_to_keep_after_gating)
  number_of_events_after_gating <-  number_of_events(data = fcs_data, file_names = file_names)
  percent_lost_each_gating[as.character(file_names),"Ce140Di"] <- number_of_events_after_gating/number_of_events_raw_data * 100 
  percent_lost_from_full_dataset[as.character(file_names),"Ce140Di"] <- number_of_events_after_gating/number_of_events_raw_data * 100 
  random_events_for_plotting <- random_events(number_of_events_after_gating)
  
  
  number_of_events_after_beads_gating <- number_of_events_after_gating
```

Since we are finished working with beads_data, remove them to save space in memory

```{r}
rm(beads_data)
```

# clean_up_data
 
 * the clean up channel names can be found by using grep of the string "Center|Offset|Width|Residual|Event|Ir191|Ir193|Pt195Di|Pt194Di"
 * a arc_sinh transformed dataset with this channels are made using the function arc_sinh_transform_selected_channels()

 
```{r}
#************************************************
  #clean_up_data
  #************************************************
  clean_up_channels <- as.character(params$name[grep("Center|Offset|Width|Residual|Event|Ir191|Ir193|Pt195Di|Pt194Di", params$name)])  #Pt194Di og 195 tilsvarer Cis
  clean_up_data <-  arc_sinh_transform_selected_channels(fcs_data = fcs_data, channels = clean_up_channels)
```

to be abel to calculate the number of events lost by this gateing we make a varialbe number_of_events_before_clean_up_gating using the function number_of_events()

```{r}
number_of_events_before_clean_up_gating <-  number_of_events(data = fcs_data, file_names = file_names)
```

# gating on Residual

the first clean up gating is on residuls. Here




```{r}
# 
# 
# 
#   #************************************************
#   #gating on Residual+ ----
#   #************************************************
#   number_of_events_before_residual_gating <-  number_of_events(data = fcs_data, file_names = file_names)
#   random_events_for_plotting <- random_events(number_of_events_before_residual_gating)
#   
#   
#   #update lower_gate_percent, upper_gate_percent
#   residual_gates <- find_gaussian_gates_second_top(data = clean_up_data, channel = "Residual", lower_gate_percent = 10, upper_gate_percent = 10)
#   density_plots <- density_plot(data = clean_up_data, "Residual", plot_title = file_names, lower_gate = residual_gates$lower_gate, upper_gate = residual_gates$upper_gate)
#   #density_plots
#   #time_signal_plots <- time_signal_plot(data = clean_up_data, random_events = random_events_for_plotting, channel = "Residual", plot_title = file_names,  lower_gate = residual_gates$lower_gate, upper_gate = residual_gates$upper_gate)
#   #time_signal_plots # to see all plots
#   #time_signal_plots[1] # to see first plot
#   
#   
#   #tiff(fs::path(outFigPath, paste0("fig2_residual_gating", ii, ".tiff")))
#   print(density_plots)
#   #dev.off()
#   
#   events_to_keep_after_gating <- events_to_keep(data = clean_up_data, channel = "Residual", lower_gate = residual_gates$lower_gates, 
#                                                 upper_gate = residual_gates$upper_gate)
#   #percent_to_keep_this_gating(kept_events = events_to_keep_after_gating, file_names = file_names)
#   
#   fcs_data <- update_data_based_on_events_to_keep(data = fcs_data, kept_events = events_to_keep_after_gating)
#   clean_up_data <- update_data_based_on_events_to_keep(data = clean_up_data, kept_events = events_to_keep_after_gating)
#   number_of_events_after_residual_gating <-  number_of_events(data = fcs_data, file_names = file_names)
#   percent_lost_from_full_dataset[as.character(file_names),"Residual"] <- number_of_events_after_residual_gating/number_of_events_raw_data * 100 #percent remaining from total
#   percent_lost_each_gating[as.character(file_names),"Residual"] <- number_of_events_after_residual_gating/number_of_events_after_beads_gating * 100 #percent remaining from bead gating
#   
#   
#   
#   #************************************************
#   #gating on Center+ ----
#   #************************************************
#   number_of_events_before_center_gating <- number_of_events_after_residual_gating
#   random_events_for_plotting <- random_events(number_of_events_before_center_gating)
#   
#   #update lower_gate_percent, upper_gate_percent
#   center_gates <- find_gaussian_gates_second_top(data = clean_up_data, channel = "Center", lower_gate_percent = 10, upper_gate_percent = 10)
#   density_plots <- density_plot(data = clean_up_data, "Center", plot_title = file_names, lower_gate = center_gates$lower_gate, upper_gate = center_gates$upper_gate)
#   #density_plots
#   #time_signal_plots <- time_signal_plot(data = clean_up_data, random_events = random_events_for_plotting, channel = "Center", plot_title = file_names, lower_gate = center_gates$lower_gate, upper_gate = center_gates$upper_gate)
#   #time_signal_plots # to see all plots
#   #time_signal_plots[1] # to see first plot
#   
#   
#   tiff(fs::path(outFigPath, paste0("fig3_event_gating", ii, ".tiff")))
#   print(density_plots)
#   dev.off()
#   events_to_keep_after_center_gating <- events_to_keep(data = clean_up_data, channel = "Center", lower_gate = center_gates$lower_gate, upper_gate = center_gates$upper_gate)
#   #percent_to_keep_this_gating(kept_events = events_to_keep_after_center_gating, file_names = file_names)
#   
#   
#   fcs_data <- update_data_based_on_events_to_keep(data = fcs_data, kept_events = events_to_keep_after_center_gating)
#   clean_up_data <- update_data_based_on_events_to_keep(data = clean_up_data, kept_events = events_to_keep_after_center_gating)
#   number_of_events_after_center_gating <-  number_of_events(data = fcs_data, file_names = file_names)
#   percent_lost_from_full_dataset[as.character(file_names),"Center"] <- number_of_events_after_center_gating/number_of_events_raw_data * 100 #percent remaining from total
#   percent_lost_each_gating[as.character(file_names),"Center"] <-number_of_events_after_center_gating/number_of_events_after_residual_gating * 100 #percent remaining from residual gating
#   
#   
#   
#   #************************************************
#   #gating on Offset+ ----
#   #************************************************
#   number_of_events_before_offset_gating <-  number_of_events_after_center_gating
#   random_events_for_plotting <- random_events(number_of_events_before_offset_gating)
#   
#   
#   #update lower_gate_percent, upper_gate_percent
#   offset_gates <- find_gaussian_gates_second_top(data = clean_up_data, channel = "Offset", lower_gate_percent = 7, upper_gate_percent = 7)
#   density_plots <- density_plot(data = clean_up_data, "Offset", plot_title = file_names, lower_gate = offset_gates$lower_gate, upper_gate = offset_gates$upper_gate)
#   #density_plots
#   #time_signal_plots <- time_signal_plot(data = clean_up_data, random_events = random_events_for_plotting, channel = "Offset", plot_title = file_names, lower_gate = offset_gates$lower_gate, upper_gate = offset_gates$upper_gate)
#   #time_signal_plots # to see all plots
#   #time_signal_plots[1] # to see first plot
#   
#   
#   tiff(fs::path(outFigPath, paste0("fig4_offset_gating", ii, ".tiff")))
#   print(density_plots)
#   dev.off()
#   events_to_keep_after_offset_gating <- events_to_keep(data = clean_up_data, channel = "Offset",  lower_gate = offset_gates$lower_gate, upper_gate = offset_gates$upper_gate)
#   #percent_to_keep_this_gating(kept_events = events_to_keep_after_offset_gating, file_names = file_names)
#   
#   
#   
#   fcs_data <- update_data_based_on_events_to_keep(data = fcs_data, kept_events = events_to_keep_after_offset_gating)
#   clean_up_data <- update_data_based_on_events_to_keep(data = clean_up_data, kept_events = events_to_keep_after_offset_gating)
#   number_of_events_after_offset_gating <- number_of_events(data = fcs_data, file_names = file_names)
#   percent_lost_from_full_dataset[as.character(file_names),"Offset"] <- number_of_events_after_offset_gating/number_of_events_raw_data * 100 #percent remaining from total
#   percent_lost_each_gating[as.character(file_names),"Offset"] <- number_of_events_after_offset_gating/number_of_events_after_center_gating * 100 #percent remaining from  center gating
#   
#   #************************************************
#   #gating on Width+ ----
#   #************************************************
#   number_of_events_before_width_gating <-  number_of_events(data = fcs_data, file_names = file_names)
#   random_events_for_plotting <- random_events(number_of_events_before_width_gating)
#   
#   
#   
#   #update lower_gate_percent, upper_gate_percent
#   width_gates <- find_gaussian_gates_highest_top(data = clean_up_data, channel = "Width", lower_gate_percent = 2.5, upper_gate_percent = 4)
#   
#   density_plots <- density_plot(data = clean_up_data, "Width", plot_title = file_names, lower_gate = width_gates$lower_gate, upper_gate = width_gates$upper_gate)
#   #density_plots
#   
#   
#   tiff(fs::path(outFigPath, paste0("fig5_width_gating", ii, ".tiff")))
#   print(density_plots)
#   dev.off()
#   #time_signal_plots <- time_signal_plot(data = clean_up_data, random_events = random_events_for_plotting, channel = "Width", plot_title = file_names, lower_gate = width_gates$lower_gate, upper_gate = width_gates$upper_gate)
#   #time_signal_plots # to see all plots
#   #time_signal_plots[1] # to see first plot
#   
#   
#   events_to_keep_after_width_gating <- events_to_keep(data = clean_up_data, channel = "Width", lower_gate = width_gates$lower_gate, upper_gate = width_gates$upper_gate)
#   #percent_to_keep_this_gating(kept_events = events_to_keep_after_width_gating, file_names = file_names)
#   
#   
#   fcs_data <- update_data_based_on_events_to_keep(data = fcs_data, kept_events = events_to_keep_after_width_gating)
#   clean_up_data <- update_data_based_on_events_to_keep(data = clean_up_data, kept_events = events_to_keep_after_width_gating)
#   number_of_events_after_width_gating <- number_of_events(data = fcs_data, file_names = file_names)
#   percent_lost_from_full_dataset[as.character(file_names),"Width"] <- number_of_events_after_width_gating/number_of_events_raw_data * 100 #percent remaining from total
#   percent_lost_each_gating[as.character(file_names),"Width"] <- number_of_events_after_width_gating/number_of_events_after_offset_gating * 100 #percent remaining from offset gating
#   
#   
#   
#   
#   #************************************************
#   #gating on Event+ ----
#   #************************************************
#   number_of_events_before_event_gating <-  number_of_events(data = fcs_data, file_names = file_names)
#   random_events_for_plotting <- random_events(number_of_events_before_event_gating)
#   
#   #update lower_gate_percent, upper_gate_percent
#   EventGates <- find_gaussian_gates_second_top(data = clean_up_data, channel = "Event_length", lower_gate_percent = 20, upper_gate_percent = 20)
#   density_plots <- density_plot(data = clean_up_data, "Event_length", plot_title = file_names, lower_gate = EventGates$lower_gate, upper_gate = EventGates$upper_gate)
#   #density_plots
#   
#   
#   tiff(fs::path(outFigPath, paste0("fig6_event_gating", ii, ".tiff")))
#   print(density_plots)
#   dev.off()
#   #time_signal_plots <- time_signal_plot(data =  clean_up_data, random_events = random_events_for_plotting, channel = "Event_length", plot_title = file_names, lower_gate = EventGates$lower_gate, upper_gate = EventGates$upper_gate)
#   #time_signal_plots # to see all plots
#   #time_signal_plots[1] # to see first plot
#   
#   events_to_keep_after_event_gating <- events_to_keep(data = clean_up_data, channel = "Event_length",  upper_gate = EventGates$upper_gate)
#   #percent_to_keep_this_gating(kept_events = events_to_keep_after_event_gating, file_names = file_names)
#   
#   
#   
#   fcs_data <- update_data_based_on_events_to_keep(data = fcs_data, kept_events = events_to_keep_after_event_gating)
#   clean_up_data <- update_data_based_on_events_to_keep(data = clean_up_data, kept_events = events_to_keep_after_event_gating)
#   number_of_events_after_event_gating <-  number_of_events(data = fcs_data, file_names = file_names)
#   percent_lost_from_full_dataset[as.character(file_names),"Event_length"] <- number_of_events_after_event_gating/number_of_events_raw_data * 100 #percent remaining from total
#   percent_lost_each_gating[as.character(file_names),"Event_length"] <- number_of_events_after_event_gating/number_of_events_after_width_gating * 100 #percent remaining from width gating
#   
#   
#   
#   
#   #************************************************
#   #gating on Live Dead----  Cis, 
#   #************************************************
#   number_of_events_before_cis_gating <-  number_of_events(data = fcs_data, file_names = file_names)
#   random_events_for_plotting <- random_events(number_of_events_before_cis_gating)
#   
#   
#   #update lower_gate_percent, upper_gate_percent
#   cis_gates <- find_gaussian_gates_second_top(data = clean_up_data, channel = "Pt194Di", lower_gate_percent = 5, upper_gate_percent = 30)
#   if(!is.na(tvungetLavereCISgate)){
#     cis_gates$lower_gates <- rep(tvungetLavereCISgate, length(cis_gates$lower_gates))
#   }
#   
#   density_plots <- density_plot(data = clean_up_data, "Pt194Di", plot_title = file_names, lower_gate = cis_gates$lower_gate, upper_gate = cis_gates$upper_gate)
#   #density_plots
#   
#   
#   tiff(fs::path(outFigPath, paste0("fig7_cis_gating", ii, ".tiff")))
#   print(density_plots)
#   dev.off()
#   #time_signal_plots <- time_signal_plot(data = clean_up_data, random_events = random_events_for_plotting, channel = "Pt194Di", plot_title = file_names, lower_gate = cis_gates$lower_gate, upper_gate = cis_gates$upper_gate)
#   #time_signal_plots # to see all plots
#   #time_signal_plots[1] # to see first plot
#   #time_signal_plots <- time_signal_plot(data = clean_up_data, random_events = random_events_for_plotting, channel = "Pt194Di", plot_title = file_names, lower_gate = NA, upper_gate = NA)
#   #time_signal_plots # to see all plots
#   events_to_keep_after_cis_gating <- events_to_keep(data = clean_up_data, channel = "Pt194Di",  lower_gate = cis_gates$lower_gate,  upper_gate = cis_gates$upper_gate)
#   #percent_to_keep_this_gating(kept_events = events_to_keep_after_cis_gating, file_names = file_names)
#   
#   
#   
#   fcs_data <- update_data_based_on_events_to_keep(data = fcs_data, kept_events = events_to_keep_after_cis_gating)
#   clean_up_data <- update_data_based_on_events_to_keep(data = clean_up_data, kept_events = events_to_keep_after_cis_gating)
#   number_of_events_after_cis_gating <-  number_of_events(data = fcs_data, file_names = file_names)
#   percent_lost_from_full_dataset[as.character(file_names),"Pt194Di"] <- number_of_events_after_cis_gating/number_of_events_raw_data * 100 #percent remaining from total
#   percent_lost_each_gating[as.character(file_names),"Pt194Di"] <- number_of_events_after_cis_gating/number_of_events_after_event_gating * 100 #percent remaining from event gating
#   
#   
#   
#   #************************************************
#   #gating on DNA1, Ir191Di+ ----
#   #************************************************
#   number_of_events_before_Ir191Di_gating <-  number_of_events(fcs_data, file_names = file_names)
#   random_events_for_plotting <- random_events(number_of_events_before_Ir191Di_gating)
#   
#   #update lower_gate_percent, upper_gate_percent
#   # Ir191di_gates <- find_gaussian_gates_second_top(data = clean_up_data, channel = "Ir191Di", lower_gate_percent = NA, upper_gate_percent = NA, perc_included = 0.99995, main_top_to_left = F)
#   # 
#   # density_plots <- density_plot(data = clean_up_data, "Ir191Di", plot_title = file_names, lower_gate = Ir191di_gates$lower_gate, upper_gate = Ir191di_gates$upper_gate)
#   #density_plots
#   
#   
#   Ir191di_gates <- find_gaussian_gates_second_top(data = clean_up_data, channel = "Ir191Di", lower_gate_percent = NA, upper_gate_percent = NA, perc_included = 0.99, main_top_to_left = F)
#   
#   density_plots <- density_plot(data = clean_up_data, "Ir191Di", plot_title = file_names, lower_gate = Ir191di_gates$lower_gate, upper_gate = Ir191di_gates$upper_gate)
#   
#   
#   tiff(fs::path(outFigPath, paste0("fig8_Ir191_gating", ii, ".tiff")))
#   print(density_plots)
#   dev.off()
#   #time_signal_plots <- time_signal_plot(data = clean_up_data, random_events = random_events_for_plotting, channel = "Ir191Di", plot_title = file_names, lower_gate = Ir191di_gates$lower_gate, upper_gate = Ir191di_gates$upper_gate)
#   #time_signal_plots # to see all plots
#   #time_signal_plots[1] # to see first plot
#   
#   events_to_keep_after_Ir191Di_gating <- events_to_keep(data = clean_up_data, channel = "Ir191Di",  lower_gate = Ir191di_gates$lower_gate,  upper_gate = Ir191di_gates$upper_gate)
#   #percent_to_keep_this_gating(kept_events = events_to_keep_after_Ir191Di_gating, file_names = file_names)
#   
#   
#   fcs_data <- update_data_based_on_events_to_keep(data = fcs_data, kept_events = events_to_keep_after_Ir191Di_gating)
#   clean_up_data <- update_data_based_on_events_to_keep(data = clean_up_data, kept_events = events_to_keep_after_Ir191Di_gating)
#   number_of_events_after_Ir191Di_gating <-  number_of_events(data = fcs_data, file_names = file_names)
#   percent_lost_from_full_dataset[as.character(file_names),"Ir191Di"] <- number_of_events_after_Ir191Di_gating/number_of_events_raw_data * 100 #percent remaining from total
#   percent_lost_each_gating[as.character(file_names),"Ir191Di"] <- number_of_events_after_Ir191Di_gating/number_of_events_after_cis_gating * 100 #percent remaining from cis gating
#   
#   
#   
#   #************************************************
#   #gating on DNA1, Ir193Di+ ----
#   #************************************************
#   number_of_events_before_Ir193Di_gating <-  number_of_events(fcs_data, file_names = file_names)
#   random_events_for_plotting <- random_events(number_of_events_before_Ir193Di_gating)
#   
#   #update lower_gate_percent, upper_gate_percent
#   Ir193di_gates <- find_gaussian_gates_highest_top(data = clean_up_data, channel = "Ir193Di", lower_gate_percent = 15, upper_gate_percent = 5)
#   
#   
#   density_plots <- density_plot(data = clean_up_data, "Ir193Di", plot_title = file_names, lower_gate = Ir193di_gates$lower_gate, upper_gate = Ir193di_gates$upper_gate)
#   
#   
#   tiff(fs::path(outFigPath, paste0("fig9_Ir193_gating", ii, ".tiff")))
#   print(density_plots)
#   dev.off()#density_plots
#   #time_signal_plots <- time_signal_plot(data = clean_up_data, random_events = random_events_for_plotting, channel = "Ir193Di", plot_title = file_names, lower_gate = Ir193di_gates$lower_gate, upper_gate = Ir193di_gates$upper_gate)
#   #time_signal_plots # to see all plots
#   #time_signal_plots[1] # to see first plot
#   
#   events_to_keep_after_Ir193Di_gating <- events_to_keep(data = clean_up_data, channel = "Ir193Di",  lower_gate = Ir193di_gates$lower_gate,  upper_gate = Ir193di_gates$upper_gate)
#   #percent_to_keep_this_gating(kept_events = events_to_keep_after_Ir193Di_gating, file_names = file_names)
#   
#   
#   fcs_data <- update_data_based_on_events_to_keep(data = fcs_data, kept_events = events_to_keep_after_Ir193Di_gating)
#   clean_up_data <- update_data_based_on_events_to_keep(data = clean_up_data, kept_events = events_to_keep_after_Ir193Di_gating)
#   number_of_events_after_Ir193Di_gating <-  number_of_events(data = fcs_data, file_names = file_names)
#   percent_lost_from_full_dataset[as.character(file_names),"Ir193Di"] <- number_of_events_after_Ir193Di_gating/number_of_events_raw_data * 100 #percent remaining from total
#   percent_lost_each_gating[as.character(file_names),"Ir193Di"] <- number_of_events_after_Ir193Di_gating/number_of_events_after_cis_gating * 100 #percent remaining from cis gating
#   
#   
#   #************************************************
#   #save fcs_data ----
#   #************************************************
#   
#   flowCore::write.flowSet(fcs_data, outdir = outDataPath, filename = as.character(file_names))
#   print(".")
# }
# 
# write.csv2(percent_loss_each_gating, fs::path(out_result, "percent_kept_each_gatingnr2.csv"))
# write.csv2(percent_loss_from_full_dataset, fs::path(out_result, "percent_kept_from_full_datasetnr2.csv"))

#proc.time() - ptm
```

