---
title: "Introduction to FlowSOM analyse"
author: "Anja Bråthen Kristoffersen"
date: "23 March 2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document describes the different lines of the file "runFlowSOM_analyse.R". To perform your own flowSOM analysis save the files "runFlowSOM_analyse.R" and  "FlowSOM_analyse.R" from github https://github.com/folkehelseinstituttet/cytof give it a new name and change the necessary paths and values throughout the document to use it on your own data. The file "FlowSOM_analyse.R" is sourced from the file "runFlowSOM_analyse.R" see also the explanatory file called "Introduction to FlowSOM_analyse.R". These documents only make a clustering of your data and some additional files that later will be used when analysing the flowSOM data together with metadata, examples of analysing of the clustering data together with metadata can be found in XXX. 


This document start with removing everything that you may already have in R to make sure you start with no previous data. The removing is done by using the functions rm() and ls(). The function ls() list all objects in R while rm() remove everything in the list included

```{r, eval=F, echo=T}
rm(list = ls())
```

Then some librarys are loaded. If you do not have these libraries installed you have to install them by using the function install.packages("LIBRARY NAME")

```{r, eval=F, echo=T}
library(ComplexHeatmap)
library(gridExtra)
library(ggplot2)
library(grid)
```

this chunk is optinal. If you want the clustering of the median signal to be clustered on markeres the order will not be used. But if you do not want to cluster the markers this will give the order you want the markers to be shown in.  

```{r, eval=F, echo=T}
orderPanel1 <- c("X89Y_CD45", "X116Cd_CD3", "X145Nd_CD4", "X113Cd_CD8", "X111Cd_CD19", "X152Sm_TCRgd", "X166Er_TCRVa7.2", "X149Sm_CD25", 
       "X158Gd_CD27", "X160Gd_CD28", "X143Nd_CD127", "X172Yb_CD38", "X167Er_CCR7", "X155Gd_CD45RA", "X114Cd_HLADR", "X146Nd_IgD", "X159Tb_IgG",
       "X156Gd_CXCR3", "X153Eu_CCR4", "X141Pr_CCR6", "X171Yb_CXCR5", "X168Er_ICOS", "X142Nd_KLRG1", "X150Nd_CD134_OX40", "X154Sm_TIGIT",
       "X161Dy_CD160", "X164Dy_CD161", "X162Dy_CD95", "X163Dy_CRTH2", "X165Ho_CD85j", "X169Tm_NKG2A", "X173Yb_CD141", "X174Yb_CD279_PD.1",
       "X175Lu_CD14", "X148Nd_CD16", "X176Yb_CD56", "X106Cd_CD57", "X209Bi_CD11b", "X147Sm_CD11c", "X112Cd_CD5", "X144Nd_CD15", "X170Er_CD169", "X151Eu_CD123")

orderPanel2 <- c("X89Y_CD45", "X116Cd_CD3", "X113Cd_CD8", "X145Nd_CD4",  "X111Cd_CD19",  "X153Eu_CXCR5.CD185", 
       "X114Cd_HLADR", "X175Lu_CD14", "X209Bi_CD16",  "X176Yb_CD56", "X106Cd_CD57", "X158Gd_CD27",
       "X160Gd_CD28", "X169Tm_CD25", "X172Yb_CD38", "X112Cd_CD44", "X143Nd_CD127.IL7Ra",  "X167Er_CCR7.CD197",
       "X155Gd_CD45RA", "X162Dy_FoxP3", "X170Er_CTLA.4.CD152", "X163Dy_CD33",  "X154Sm_CD272.BTLA", "X110Cd_CD107a",
       "X161Dy_IL.17A", "X166Er_IL.10", "X164Dy_Perforin", "X171Yb_GranzymeB", "X148Nd_CD274.PD.L1",  
       "X173Yb_CD273.PD.L2", "X159Tb_GM.CSF",   "X168Er_CD154.CD40L", "X174Yb_CD279.PD1", "X141Pr_CD223.LAG3", 
       "X147Sm_TIM.3",  "X151Eu_CD137.4.1BB",  "X165Ho_IFNg", "X142Nd_IL.1b", "X156Gd_IL.6", "X166Er_IL.10", 
       "X152Sm_TCRgd", "X151Eu_CD137.4.1BB", "X149Sm_IL.12p70", "X150Nd_MIP.1b")
```


then the scriptPath has to be updated to where you have saved your script. Read the data you want to analyse. and deside which files you want to use in this analyses.


```{r, eval=F, echo=T}
scriptPath <- fs::path("H:", "git", "cytof")
#scriptPath <- fs::path("C:", "CyToF data", "fra github")

source(fs::path(scriptPath, "readDataToAnalysePanel2.R"))

tamed <- as.character(dInfo$filenames)
tamed <- tamed[!grepl("FHI005", tamed)] #denne oppfører seg helt rart og tas ut av analysen.
tamed <- tamed[!grepl("Ref1", tamed)] #denne oppfører seg helt rart og tas ut av analysen.
tamed
```

ks is then assigned to be a vector with the number of clusteres you want to analyse. Here you could use one or more numbers. xdim and ydim tells how many nodes that flowSOM will use that later will be clustered to k clusters defined in ks. By setting xdim and ydim to 14 there will be 14*14 = 196 nodes, so the ks has to be less than 196 and probably as a rule at most half of the number of nodes. the variable n_per_file tells how many cells to use per file.

```{r, eval=F, echo=T}
ks <- c(40, 50, 60)
n_per_file <- 25000
xdim <- 14
ydim <- 14
```

Here the path to where you want to save the data is assigned, this has to be changed according to your data. Together with a seed that ensure that you get the same result if you run the script one more time. The seed has to be changed to randomly draw an other set of cells to analyse. If selectedEvents are TRUE channel has to be given as well. selectedEvents tells if you will draw from all cells or only from a selected number of cells here by using channel equal to the rowname in posNeg (the list for each file of 0 (negative), 1 (positive/low), 2 (high))

```{r, eval=F, echo=T}
utSti <- fs::path("F:", "Forskningsprosjekter", "PDB 2794 - Immune responses aga_", "Forskningsfiler", "JOBO", "CyTOF", "Analyse i R OUS", "Resultat_Panel_2_ALLE")
seed <- 12345 #nb må endres vil man vil gjøre et annet uttrekk
selectedEvents <- FALSE
channel <- NULL
o <- orderPanel2
column_cluster <- FALSE

source(fs::path(scriptPath, "FlowSOM_analyse.R"))



for(j in 1:length(posNeg)){
  posNeg[[j]]$CD3CD45CD4 <- posNeg[[j]]$CD3 & posNeg[[j]]$CD45 & posNeg[[j]]$CD4
}

utSti <- fs::path("F:", "Forskningsprosjekter", "PDB 2794 - Immune responses aga_", "Forskningsfiler", "JOBO", "CyTOF", "Analyse i R OUS", "Resulta_Panel_2_ALLE_CD4_test")
seed <- 45903#nb må endres vil man vil gjøre et annet uttrekk
selectedEvents <- TRUE
channel <- "CD3CD45CD4"
n_per_file <- 4250
ks <- c(15, 20, 25, 30, 35, 40)
xdim <- 10
ydim <- 10
o <- orderPanel2
column_cluster <- FALSE

source(fs::path(scriptPath, "FlowSOM_analyse.R"))





for(j in 1:length(posNeg)){
  posNeg[[j]]$CD3CD45CD8 <- posNeg[[j]]$CD3 & posNeg[[j]]$CD45 & posNeg[[j]]$CD8
}

utSti <- fs::path("F:", "Forskningsprosjekter", "PDB 2794 - Immune responses aga_", "Forskningsfiler", "JOBO", "CyTOF", "Analyse i R OUS", "Resulta_Panel_2_ALLE_CD8")
seed <- 45902#nb må endres vil man vil gjøre et annet uttrekk
selectedEvents <- TRUE
channel <- "CD3CD45CD8"
n_per_file <- 4000
ks <- c(15, 20, 25, 30, 35, 40)
xdim <- 10
ydim <- 10
o <- orderPanel2
column_cluster <- FALSE


source(fs::path(scriptPath, "FlowSOM_analyse.R"))

#panel 1

source(fs::path(scriptPath, "readDataToAnalyse.R"))

tamed <- as.character(dInfo$filenames)
tamed <- tamed[!grepl("FHI005", tamed)] #denne oppfører seg helt rart og tas ut av analysen.
tamed <- tamed[!grepl("Ref1", tamed)] #denne oppfører seg helt rart og tas ut av analysen.
tamed

ks <- c(40, 50, 60)
n_per_file <- 25000
xdim <- 14
ydim <- 14


utSti <- fs::path("F:", "Forskningsprosjekter", "PDB 2794 - Immune responses aga_", "Forskningsfiler", "JOBO", "CyTOF", "Analyse i R OUS", "Resulta_Panel_1_ALLE")
seed <- 1245 #nb må endres vil man vil gjøre et annet uttrekk
selectedEvents <- FALSE
channel <- NULL
o <- orderPanel1
column_cluster <- FALSE

source(fs::path(scriptPath, "FlowSOM_analyse.R"))



for(j in 1:length(posNeg)){
  posNeg[[j]]$CD3CD45CD4 <- posNeg[[j]]$CD3 & posNeg[[j]]$CD45 & posNeg[[j]]$CD4
}

utSti <- fs::path("F:", "Forskningsprosjekter", "PDB 2794 - Immune responses aga_", "Forskningsfiler", "JOBO", "CyTOF", "Analyse i R OUS", "Resulta_Panel_1_ALLE_CD4")
seed <- 4503#nb må endres vil man vil gjøre et annet uttrekk
selectedEvents <- TRUE
channel <- "CD3CD45CD4"
n_per_file <- 4250
ks <- c(15, 20, 25, 30, 35, 40)
xdim <- 10
ydim <- 10
o <- orderPanel1
column_cluster <- FALSE


source(fs::path(scriptPath, "FlowSOM_analyse.R"))





for(j in 1:length(posNeg)){
  posNeg[[j]]$CD3CD45CD8 <- posNeg[[j]]$CD3 & posNeg[[j]]$CD45 & posNeg[[j]]$CD8
}

utSti <- fs::path("F:", "Forskningsprosjekter", "PDB 2794 - Immune responses aga_", "Forskningsfiler", "JOBO", "CyTOF", "Analyse i R OUS", "Resulta_Panel_1_ALLE_CD8")
seed <- 4502#nb må endres vil man vil gjøre et annet uttrekk
selectedEvents <- TRUE
channel <- "CD3CD45CD8"
n_per_file <- 4000
ks <- c(15, 20, 25, 30, 35, 40)
xdim <- 10
ydim <- 10
o <- orderPanel1
column_cluster <- FALSE


source(fs::path(scriptPath, "FlowSOM_analyse.R"))


```