---
title: "Vignette clean up gateing"
author: "Anja Bråthen Kristoffersen"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

last updated

```{r}
date()
```


# To get started
 
  This part can be done by running the first lines in the script "run.R". This is only an explanation of what these lines does. 


  * make a folder where you want to save the result from you analysis. In this example I have called it 

```{r, include=TRUE, message= FALSE, comment= FALSE}
path0 <- fs::path("C:", "Cytof data", "test")  
path0
```

 * make a folder "script" inside this folder, and copy all files from GitHub in here
 * then source the file "function.R

```{r, include=TRUE, message= FALSE, comment= FALSE}
source(fs::path(path0, "script", "functions.R"))
```

 * and run the next line which will make a vector of folder names that also are created in path0. The function make_project_folders() use the packages org form fhi. See install libraries 

```{r, include=TRUE, message= FALSE, comment= FALSE}
paths <- make_project_folders(path0)
rm(path0)
```

this is the folders you have made 

```{r}
paths
```

 * place all your raw data in the folder "raw data"

