---
title: "Vignette run cytof analysis"
author: "Anja Bråthen Kristoffersen"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

last updated

```{r}
date()
```

# To get started
 
  This explains the lines in the script "run.R". This is only an explanation of what these lines does. 
  
  * First R has to be told which librarys to use. If you have not installed the packages, see "Example-install-packages-in-R.docx".


```{r, eval = F}
library(ComplexHeatmap)
library(gridExtra)
library(ggplot2)
library(grid)
library(flowCore)
library(FlowSOM)
library(org)
library(data.table)
library(fs)
library(betareg)
```




  * make a folder where you want to save the result from you analysis. In this example I have called it 

```{r, include=TRUE, message= FALSE, comment= FALSE}
path0 <- fs::path("C:", "Cytof data", "test")  
path0
```

 * make a folder "script" inside this folder, and copy all files from GitHub in here
 * then source the file "function.R

```{r, include=TRUE, message= FALSE, comment= FALSE}
source(fs::path(path0, "script", "functions.R"))
```

 * and run the next line which will make a vector of folder names that also are created in path0. 

```{r, include=TRUE, message= FALSE, comment= FALSE}
paths <- make_project_folders(path0)
rm(path0)
```

The function make_project_folders() use the packages org form FHI. If you have problems running this, you could rather make the folder structure as below. This structure is needed for the rest of the script to run properly.  

```{r}
paths
```

 * place all your raw data in the folder "raw data".
 
 * then before the analysis can be started the function plotSignal() has to be updated. Here I have made an example with 15 files (and one commented away with 66 files), change this function so that it takes as many files as your dataset have. 


```{r, include=FALSE, message= FALSE, comment= FALSE}
#THIS HAS TO BE CHANGED ACCORDING TO NUMBER OF FILES IN YOUR DATASET.
#the plot_list have to include as many memberes as there are files in your dataset, also remember to update ncol and nrow.
plotSignal <- function(plot_list){
  g <- gridExtra::grid.arrange(plot_list[[1]], plot_list[[2]],
                               plot_list[[3]], plot_list[[4]], plot_list[[5]],
                               plot_list[[6]],  plot_list[[7]], plot_list[[8]],
                               plot_list[[9]], plot_list[[10]], plot_list[[11]],
                               plot_list[[12]], plot_list[[13]], plot_list[[14]],
                               plot_list[[15]], #plot_list[[16]], plot_list[[17]],
                               # plot_list[[18]], plot_list[[19]], plot_list[[20]],
                               # plot_list[[21]], plot_list[[22]], plot_list[[23]],
                               # plot_list[[24]], plot_list[[25]], plot_list[[26]],
                               # plot_list[[27]], plot_list[[28]], plot_list[[29]],
                               # plot_list[[30]], plot_list[[31]], plot_list[[32]],
                               # plot_list[[33]], plot_list[[34]], plot_list[[35]],
                               # plot_list[[36]], plot_list[[37]], plot_list[[38]],
                               # plot_list[[39]], plot_list[[40]], plot_list[[41]],
                               # plot_list[[42]], plot_list[[43]], plot_list[[44]],
                               # plot_list[[45]], plot_list[[46]], plot_list[[47]],
                               # plot_list[[48]], plot_list[[49]], plot_list[[50]],
                               # plot_list[[51]], plot_list[[52]], plot_list[[53]],
                               # plot_list[[54]], plot_list[[55]], plot_list[[56]],
                               # plot_list[[57]], plot_list[[58]], plot_list[[59]],
                               # plot_list[[60]], plot_list[[61]], plot_list[[62]],
                               # plot_list[[63]], plot_list[[64]], plot_list[[65]],
                               # plot_list[[66]],  ncol = 11, nrow = 6)
                               ncol = 5,nrow = 3)
  return(g)
}
```

###################
# CLEAN Up 
###################

You are now ready to start the clean up of your data. Run the lines below to get a first cleanup version. 

When you are happy with your cleanup you only have to run it once, that is why I have written it within an if-sentence. 


```{r, eval = F}
if(nrow(file.info(list.files(paths$clean_data_path))) == 0){
  source(fs::path(paths$script_path, "clean up gating all files.R"))
}

#
```

Check the gates produced in both density plots and time-signal-plots which you find in "...\clean_up\clean_data_fig_density" and "...\clean_up\clean_data_fig_signal". If something has to be changed, and most likely it will, you have to update  "clean up gating all files.R". How to do this you fine in "Vignette clean up gating.docx". And rerun by writing

```{r, eval = F}
source(fs::path(paths$script_path, "clean up gating all files.R"))
```

###################
# Read clean data 
###################

You have now cleaned your data and the clean data can be found in the folder "...\clean_up\clean_data". You have also made three files in  "...\clean_up\clean_data_info\". "percent_kept_from_full_dataset.csv" tells you how many percent of the full dataset you have kept after each gating, "percent_kept_each_gating.csv" how many percent of the gated dataset you have kept for each gating. The last file "marker_names_included_manual_shortnames.csv" are a table with marker names. The two first column must be kept as they are the last one can be changed. The last column are the names that will be used in plots and when you in the script refere to a marker. Please update the column "marker_short_name" with the names you want to use. Make sure that all markers in the list are markers that you want to use in your analysis. If there are too many markers you could either delete the rows manually or update the file "clean up gating all files.R" that make the script. If you make any update to the file "marker_names_included_manual_shortnames.csv" remember that this is produced again each time you run "marker_names_included_manual_shortnames.csv", so you might also want to save a copy with your changes. 



```{r, eval = F}
files_to_open <- "all" 
```

This will include all files in the clean data folder. If you want to do the analysis with only some files this can be changed to a vector of filenames to use.

```{r, eval = F}
rownames(file.info(list.files(paths$clean_data_path)))
```

give you a list of all files in the clean data folder, only .fcs files are read
 
 * to read the data and the updated file with marker short names write:
 
```{r, eval = F}
fcs_data_with_info <- read_data_from_folder(data_path = paths$clean_data_path, 
                                            files_to_open = files_to_open)
fcs_data <- fcs_data_with_info$fcs_data
file_names <- fcs_data_with_info$file_names
rm(fcs_data_with_info)

params_fcs <- get_params_fcs_data(fcs_data[[1]])

marker_info <- read.csv2(fs::path(paths$clean_data_info_path, "marker_names_included_manual_shortnames.csv"), stringsAsFactors = FALSE)
```

The variable fcs_data now contains a list of matrixes with your data,  while file_names are a vector of names in the same order as the data are stored in fcs_data. 

###################
# Marker gating 
###################

This section can be skipped if manual gates for the markers are not of interest, but it is needed for some markers if selected events should be used in further analysis. Once the manual gating is as you like it, it is not necessary to rerun. Hence the if sentence. If you still want to rerun just run the line within the if sentence.

The example only include gating for CD3, CD4, CD8 and CD45, all gates should be adjusted inside the file "marker gating all files.R" see "Vignette marker gating.docx" for explanation on how to make changes in the file. You might also change the gates to be equal for all files. 

```{r, eval = F}
if(nrow(file.info(list.files(paths$marker_gating_results_path)) ) == 0){
  source(fs::path(paths$script_path, "marker gating all files.R"))
}
```

This produce a list of matrices saved in "...\clean_up\marker_gating\resultsposNeg.rds" that are in the same size as fcs_data but which contains 0 (negative), 1 (positive or low positive) and 2 (high positive) for each event and each marker. The same folder also contains "mean_gates.csv" which is a table of the mean gates for those markers gated, the other markers have gotten the gate NA.  


###################
# Run FlowSOM 
###################

To do a flowSOM analyse, and make some plots associated to this analysis run the function run_flowSOM(). This function takes many parameters:

 * fcs_data: list of matrices of clean data, not transformed, as already made earlier in this script
 * file_names:  vector of the file_names in fcs_data in same order as the data, as already made earlier in this script
 * included_files: default "all", could be changed to only some files, if not included alle files are used
 * n_per_file: number of events to include per file, I here used 15000
 * included_markers: default "all", could manually be changed with a list of markers to use in analysis. The function  get_marker_name_from_marker_short_name() might be used to uptain proper marker_name
 * transformation: default "arc_sinh", by now only option. 
 * scaling_flowSOM: default TRUE,  scaling is recommanded, but will mean that all markers count as much regardless of signal size
 * k_s: number of clusters, k_s has to be a number or a vector of numbers.
 * xdim:  default 10, xdim * ydim gives the number of nodes that FlowSOM will combined to k_s clusters. With many clusters 10 might be too small (should have atleast 4 times as many nodes as clusters)
 * ydim: default 10, preferably equal to xdim
 * resultpath: recomanded paths$clean_data_flowSOM_results_path or if many different analyses are done folder within this folder
 * seed: random number that ensure that when the same seed is used the same events are chosen, and hence the same result are uptained next time the exactly same analysis are done
 * selectedEvents: default "all", if not "all" has to be the same name as a coloum-name in the matrices in posNeg, see example below
 * posNeg default NULL, since posNeg only are needed if selectedEvents are less than all events
 * make_heatmap default TRUE, need library ComplexHeatmap to make a heatmap. 
 * heatmap_cluster_column default FALSE, if FALSE the column (file_names) are not clustered and will end up in the same order as file_names are entered.


```{r, eval = F}
run_flowSOM(fcs_data = fcs_data, 
            file_names = file_names,
            included_files = "all", 
            n_per_file = 15000, 
            included_markers = "all", 
            transformation = "arc_sinh", 
            scaling_flowSOM = TRUE, 
            k_s = c(10, 15, 20), 
            xdim = 10, 
            ydim = 10, 
            resultpath = paths$clean_data_flowSOM_results_path,  
            seed = 1234, 
            selectedEvents = "all", 
            posNeg = NULL, 
            make_heatmap = TRUE, 
            heatmap_cluster_column = FALSE 
            )
```

this should be repeated for different seeds.

###################
# Run FlowSOM  for selected events 
###################

To do this marker gating has to be performed and a posNeg.rds produced (see earlier in document).

 * read the posNeg file from  "...\clean_up\marker_gating\results"

```{r, eval = F}
posNeg <- readRDS( fs::path(paths$marker_gating_results_path, "posNeg.rds"))
```

 * the function extra_column_posneg() will make an extra column to all matrices in the list posNeg. This column will in the example below get the name "CD45CD3CD8" which is the value of the parameter column_name. The markers in the matrices in posNeg that are used to make this new column are given as a vector in the parameter markers, here c("CD45", "CD3", "CD8"), in the end markers_value indicate which values this markers should have, since the order is essential here the vector c(1,1,12) is assigned to markers_values, which indicate that we want the events where CD45 = 1, CD3 = 1 and CD8 = either 1 or 2. In the gating done previously CD45 and CD3 are only 0 (negative) or 1 (positive) while CD8 is chosen to be gated as 0 (negative), 1 (low positive), 2 (high positive). Other combinations of markers and markers_value would give you other column, you could forinstance want to gate on only CD3 negative files

```{r, eval = F}
posNeg <- extra_column_posneg(posNeg = posNeg, column_name ="CD45CD3CD8", markers = c("CD45", "CD3", "CD8"), markers_values = c(1,1,12))
```

 * Other combinations of markers and markers_value would give you other column, you could forinstance want to gate on only CD3 negative files. The syntac would then be
 
```{r, eval = F}
posNeg <- extra_column_posneg(posNeg = posNeg, column_name ="CD3neg", markers = c("CD3"), markers_values = c(0))
```


We are then ready to run flowSOM on only those events that are positive for CD45, CD3 and CD8. This is done by calling run_flowSOM with the same parameters as above, except the parameter posNeg has to be equal posNeg and the parameter selectedEvents equal to the new column name in posNeg, here "CD45CD3CD8". You may also have to change n_per_file so that you in all (or atleast most) of the files have enough events to randomly choose n_per_file evenst from. It is also recomanded to save this to another folder.

The function number_of_positive_events() gives you the number of events that are positive for marker i (either 1 or 2)


```{r, eval = F}
number_of_positive_events(posNeg = posNeg, marker = "CD45CD3CD8")
```


```{r, eval = F}
run_flowSOM(fcs_data = fcs_data,
            file_names = file_names,
            included_files = file_names, 
            n_per_file = 5000,
            included_markers = marker_info$marker_name,
            transformation = "arc_sinh",
            scaling_flowSOM = TRUE, 
            k_s = c(10, 15, 20), 
            xdim = 10,
            ydim = 10, 
            resultpath = fs::path(paths$clean_data_flowSOM_results_path,  "CD45CD3CD8"),  
            seed = 2134, 
            selectedEvents = "CD45CD3CD8", 
            posNeg = posNeg, 
            make_heatmap = TRUE, 
            heatmap_cluster_column = FALSE 
)
```

###################
# Markerplot 
###################

```{r, eval = F}

resultpath <- paths$clean_data_flowSOM_results_path  # path to where the q5, q10, ... files from run_flowSOM is saved.
# if gates included in markerplots the csv file has to be made, colnames "X"	"low"	"high"
# X has to be equal column marker_short_name in ...clean_up\clean_data_info\marker_names_included_manual_shortnames.csv
# low is first gate, high second gate, both can be NA
# gates <- read.csv2(fs::path(paths$clean_data_posNeg_path, "mean_gates.csv"), stringsAsFactors = FALSE)

k <- 10
seed <- 1234
selectedEvents <- "all"
highlight_cluster <- 4
order_marker_shortname <- marker_info$marker_short_name
gates <- NULL

tiff(fs::path(resultpath, paste0("markerplot_k_", k,"_seed", seed, selectedEvents, ".tiff")), width = 1150, height = 900)
marker_plot(path = resultpath, 
            k = k, 
            seed = seed, 
            selectedEvents = selectedEvents, 
            highlight_cluster = highlight_cluster, 
            gates = gates, 
            order_marker_shortname = order_marker_shortname)
dev.off()




k <- 15
seed <- 1234
selectedEvents <- "all"
highlight_cluster <- NA
order_marker_shortname <- marker_info$marker_short_name
gates <- NULL
tiff(fs::path(resultpath, paste0("markerplot_k_", k,"_seed", seed, selectedEvents, ".tiff")), width = 1150, height = 900)
marker_plot(path = resultpath, 
            k = k, 
            seed = seed, 
            selectedEvents = selectedEvents, 
            highlight_cluster = highlight_cluster, 
            gates = gates, 
            order_marker_shortname = order_marker_shortname)
dev.off()



k <- 20
seed <- 1234
selectedEvents <- "all"
highlight_cluster <- 1:20
order_marker_shortname <- marker_info$marker_short_name
gates <- NULL
tiff(fs::path(resultpath, paste0("markerplot_k_", k,"_seed", seed, selectedEvents, ".tiff")), width = 1150, height = 900)
marker_plot(path = resultpath, 
            k = k, 
            seed = seed, 
            selectedEvents = selectedEvents, 
            highlight_cluster = highlight_cluster, 
            gates = gates, 
            order_marker_shortname = rker_info$marker_short_name)
dev.off()


#
k <- 10
seed <- 2134
selectedEvents <- "CD45CD3CD8"
highlight_cluster <- 4
order_marker_shortname <- marker_info$marker_short_name
gates <- read.csv2(fs::path(paths$marker_gating_results_path, "mean_gates.csv"))
resultpath <- fs::path(paths$clean_data_flowSOM_results_path,  "CD45CD3CD8")

tiff(fs::path(resultpath, paste0("markerplot_k_", k,"_seed", seed, selectedEvents, ".tiff")), width = 1150, height = 900)
marker_plot(path = resultpath, 
            k = k, 
            seed = seed, 
            selectedEvents = selectedEvents, 
            highlight_cluster = highlight_cluster, 
            gates = gates, 
            order_marker_shortname = order_marker_shortname)
dev.off()





```


